Noun Learning as Network Growth: A Cross-linguistic Analysis

### 2019-02-06 journal/ms.tex
second version journal

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Word Learning as Network Growth: A Cross-linguistic Analysis},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Essential manuscript parts
  \title{Word Learning as Network Growth: A Cross-linguistic Analysis}

  \shorttitle{Word Learning as Network Growth}

@@

  \abstract{Children tend to produce words earlier when they are connected to a
variety of other words along the phonological and semantic dimensions.
Though this connectivity effect has been extensively documented, little
is known about the underlying developmental mechanism. One view suggests
that learning is primarily predicted by a network growth model where
highly connected words in the child's early lexicon enable learning of
similar words. Another view suggests that learning is driven by highly
connected words in the external learning environment, instead of highly
connected words in the early internal lexicon. The present study tests
both scenarios systematically in both the phonological and semantic
domains across 10 languages. We show that external connectivity in the
learning environment drives growth in both production- and
comprehension-based vocabularies, even controlling for word frequency
and length. This pattern of findings suggests a word learning mechanism
where children harness their statistical learning abilities to detect
and learn highly connected words in the learning environment.}
  \keywords{Language understanding; audio-visual processing; word learning; speech
perception; computational modeling. \\

@@

Besides word-level properties, the structure of the lexicon (that is,
how words relate to one another) was also found to predict the Age of
Acquisition (AoA) of words. The lexical structure can be characterized
in terms of a network where each node represents a word in the
vocabulary, and each link between two nodes represents a relationship
between the corresponding pair of words (e.g., Collins \& Loftus, 1975;
Luce \& Pisoni, 1998). Previous studies have investigated early
vocabulary structure by constructing networks using a variety of
word-word relations including shared semantic features (McRae, Cree,
Seidenberg, \& McNorgan, 2005), target-cue relationships in free
association norms (Nelson, McEvoy, \& Schreiber, 1998), co-occurrence in
child directed speech (MacWhinney, 2014), and phonological relatedness
(Vitevitch, 2008). These studies have generally found that children tend
to produce words that have higher neighborhood density (i.e., high
connectivity in the network) earlier, both at the phonological and the
semantic level (Carlson, Sonderegger, \& Bane, 2014; Hills, Maouene,
Riordan, \& Smith, 2010; Hills, Maouene, Maouene, Sheya, \& Smith, 2009;
Stella, Beckage, \& Brede, 2017; Storkel, 2009).

While most studies have focused on the static properties of the lexical
network, a few have investigated the underlying developmental process.
In particular, Steyvers and Tenenbaum (2005) suggested that the observed
effects of connectivity are the consequence of how the lexical network
gets constructed in the child's mind. According to this explanation,
known as Preferential Attachment, highly connected words in the child's
lexicon tend to \enquote{attract} more words over time, in a
rich-get-richer scenario (Barabasi \& Albert, 1999). In other words,
what predicts word learning is the \emph{internal} connectivity in the
child's early lexicon. In contrast, Hills et al. (2009) suggested that
what biases the learning is not the connectivity in the child's internal
lexicon but, rather, \emph{external} connectivity in the learning
environment. They called this alternative explanation Preferential
Acquisition. For clarity of reading, we will call preferential
attachment the Internally-driven mechanism (INT), and preferential
acquisition the Externally-driven mechanism (EXT).
Figure~\ref{fig:growth} shows an illustration of both growth scenarios
with the same simplified network.

These two proposals represent two divergent ideas about the role of
lexical networks in acquisition. On the INT proposal, learning is driven
by known words with high connectivity to other known words (Figure 1,
left). Thus, the network structure is a causal factor in early word
learning, that is, children rely on the organization of their past
knowledge to determine future learning (Altvater-Mackensen \& Mani,
2013; Borovsky, Ellis, Evans, \& Elman, 2016; Chi \& Koeske, 1983;
Storkel, 2009). In contrast, on the EXT approach, learning is driven by
the connectivity of words that are not known yet (Figure 1, right).
Thus, the relevant network structure is not internally represented by
children, and the observed connectivity effect might be an epiphenomenon
of some properties of the linguistic input. For example, highly
connected words in the input could be more easily learned because of
their contextual diversity, allowing for an easier meaning
disambiguation (McMurray, Horst, \& Samuelson, 2012; Smith \& Yu, 2008;
Yurovsky \& Frank, 2015). Another reason could be that these words are
emphasized by the cargivers in their child-directed speech (Clark, 2007;
Hoff \& Naigles, 2002; Huttenlocher et al., 1991).

@@

Hills et al. (2009) investigated the growth of lexico-semantic networks
in toddlers and found that growth did not proceed according to INT as
was originally hypothesised by Steyvers and Tenenbaum (2005), but rather
according to
EXT.\footnote{Besides INT and EXT, the authors tested a third mechanism (called the lure of associates) which resembles EXT in that it is driven by the connectivity of external nodes, except that this connectivity is computed with respect to words that are known. However, EXT is the externally-driven scenario that best predicted the data in this previous work.}
This is an important finding because it suggests that learning in the
early stages is mostly driven by properties of the external input,
regardless of how past knowledge is organized. However, this work has
been done in the special case of networks that are based on 1) semantic
associations, 2) production-based vocabularies, and 3) data from
English-learning children, only. The extent to which this result depends
on the domain (e.g., semantic vs.~phonological connectivity), the
vocabulary measure (production vs.~comprehension) and culture/language
is still unclear. In this work, we test the generality of the finding
along these three dimensions.

First, we study the phonological network in addition to the semantic
network. These two networks represent different ways the mental lexicon
is structured (Beckage \& Colunga, 2016). In particular, words that are
neighbors in the semantic network (e.g., \enquote{cat}, \enquote{dog})
are not necessarily neighbors in the phonological network, and vice
versa. Does the phonological network also predict word learning?
Previous work did find an effect of words' connectivity in the
phonological network on their age of learning (Carlson et al., 2014;
Stella et al., 2017; Storkel, 2009). In other words, words learned
earlier in life tend to sound similar to many other words than a word
learned later in life. However, this finding is \emph{a priori}
compatible with both INT and EXT, and previous studies did not
explicitly compare these two mechanisms. Here, we investigate whether
phonological networks, like semantic networks, grow through EXT, or if
they rather grow via INT (Figure~\ref{fig:growth}).

Second, we study vocabularies measured using both comprehension and
production. Previous studies have found differences between these
vocabularies in terms of their content and rate of acquisition
(Benedict, 1979; Fenson et al., 1994). These differences may reflect the
fact that comprehension and production do not share the same
constraints. For instance, whereas comprehension depends on the ease
with which words are stored and accessed, production depends,
additionally, on the ease with which words are articulated, e.g.,
shorter words are produced earlier (Braginsky et al., 2016). By
investigating comprehension-based vocabularies, we assess the extent to
which the network growth mechanism captures general learning patterns
beyond the specific constraints of production.

@@

We adopted the following research strategy. We used a cross-linguistic
dataset of children's vocabulary development (Fenson et al., 1994;
Frank, Braginsky, Yurovsky, \& Marchman, 2017), and we studied the
timeline of word learning using the normative/average age of acquisition
(i.e., the age at which at least 50\% of children know a given words).
Our choice of studying the normative learning trajectory rather than the
individual trajectories was motivated by the nature of the dataset
used---which is primarily based on cross-sectional studies. Children may
vary in their individual learning trajectories, but the aggregate data
provide highly robust measures of the average learning patterns (Fenson
et al., 1994). The use of such measures has lead to important insights
on the mechanisms of word learning (Braginsky et al., 2016; J. C.
Goodman et al., 2008; Hills et al., 2010, 2009; Stella et al., 2017;
Storkel, 2009).

The paper is organized as follows. First, we describe the datasets we
used and explain how we constructed the networks. Second, we analyze
static properties of words' connectivity in these networks (correlation
with age of acquisition and shape of the distribution) and we explain
how these properties infom hypotheses about network growth. Next, we
explicitly fit the two hypothesized growth mechanisms to the data. We
investigate the extent to which the results obtained in Hills et al.
(2009) generalize to phonological networks and comprehension-based
vocabularies, and whether this generalization holds
cross-linguistically.

@@

We used data from Wordbank (Frank et al., 2017), an open repository
aggregating cross-linguistic language developmental data of the
MacArthur-Bates Communicative Development Inventory (CDI), a parent
report vocabulary checklist. Parent report is a reliable and valid
measure of children's vocabulary that allows for the cost-effective
collection of datasets large enough to test network-based models of
acquisition (Fenson et al., 1994). When filling out a CDI form,
caregivers are either invited to indicate whether their child
\enquote{understands} (comprehension) or \enquote{understands and says}
(production) each of about 400-700 words. For younger children (e.g., 8
to 18 months in the English data), both comprehension and production are
queried, whereas for older children (16 to 36 months) only production is
queried. Due to these limitations, we use data from younger children to
test comprehension and data from older children to test production.
Following previous studies (Hills et al., 2009; Storkel, 2009), we
restrict our analysis to the category of
nouns.\footnote{This choice was also forced by the fact that the great majority of association data produced by adults–and which we used to build semantic networks–were nouns.}.
Table \ref{tab:stats} gives an overview of the data we used.

@@

We constructed semantic networks for English data following the
procedure outlined in Hills et al. (2009), as follows. We used as an
index of semantic relatedness the Florida Free Association Norms (Nelson
et al., 1998). This dataset was collected by giving adult participants a
word (the cue), and asking them to write the first word that comes to
mind (the target). For example, when given the word \enquote{ball}, they
might answer with the word \enquote{game}. A pair of nodes were
connected by a directed link from the cue to the target if there was a
cue-target relationship between these nodes in the association norms.
The connectivity of a given node was characterized by its
\emph{indegree}: the number of links for which the word was the
target.\footnote{This choice was based the prior work by @hills2009 stating that analyses with both outdegrees (sum of the links where the word is the cue in a cue-target pair) and total degree  (outdegree plus indegree) led to results weaker than those calculated for indegree.}
To model growth from month to month, we constructed a different network
at each month, based on the words that have been acquired by that month.

Since the free association norms are available only in English, we used
the hand-checked translation equivalents available in Wordbank, which
allowed us to use the English association norms across 10 languages.
Using the same association data across languages does not necessarily
lead to similar networks. Indeed, though this approximation assumes that
the semantic similarity measure is universal--which is a reasonable
approximation (e.g., Youn et al., 2016), the set of words acquired by
children as well as the timeline of this acquisition can still vary from
language to language leading to possibility different learning
strategies.

@@

To construct phonological networks we first mapped the orthographic
transcription of words to their International Phonetic Alphabet (IPA)
transcriptions in each language, using the open source text-to-speech
software \textbf{\href{http://http://espeak.sourceforge.net/}{Espeak}.}
We used the Levenshtein distance (also known as edit distance) as a
measure of phonological relatedness between two nodes. The measure
counts the minimum number of operations (insertions, deletions,
substitutions) required to change one string into another.

In previous studies, two nodes were linked if they had an edit distance
of 1 (Carlson et al., 2014; Stella et al., 2017; Storkel, 2009).
However, these studies reported a contribution of phonological
connectivity to word learning when networks were built using a rich
adult vocabulary. Since the focus of the current study is on the
mechanism of growth, the networks should be based on children's early
vocabulary. The latter, however, contains very few word pairs with an
edit distance of 1. Thus, we increased the threshold from 1 to 2, that
is, two nodes were related if their edit distance was equal to 1 or
2.\footnote{We also considered the case of an edit distance of 1 as well as the continuous measure, i.e., the inverse edit distance without threshold. In both cases, the results were weaker than those obtained with a threshold of 2. We did not consider the case of a threshold larger than 2 since the pairs will become less and less phonologically related.}
The connectivity of a given node was characterized with its
\emph{degree}: the number of links it shares with other words.

@@

We start by analyzing word connectivity in the global (static) network.
We constructed this network using nouns learned by the oldest age for
which we have CDI data (e.g., in English this corresponds, in
comprehension, to the network by 18 months, and in production, to the
network by 30 months). This global network is the end-state towards
which both INT and EXT converge by the last month of learning. Moreover,
following Hills et al. (2009), we used this end-state network as a proxy
for the external connectivity in the learning environment. Below we
analyze properties of this global networks that may a priori hint at an
INT- or EXT-like growth. In order to compare various predictors on the
same data, we restrict the analysis to the subset of nouns for which we
had both semantic and phonological information in each language.

@@

Connectivity in the global network is directly related to EXT as it
represents the explicit criterion this growth scenario uses to determine
what words should be learned first (Figure \ref{fig:growth}). Therefore,
a direct consequence of an EXT-like growth scenario is a correlation
between connectivity in the global network and the age of acquisition.
This correlation is also necessary to INT, although the causality is
reversed: higher connectivity in the global network is caused by earlier
learning, not the other way around. Some words end up being highly
connected in the global network precisely because they happen to be
acquired earlier and, therefore, have a higher chance of accumulating
more links over time. Thus, the correlation between connectivity in the
end-state network and AoA can result from both EXT and INT. If there is
no such correlation, neither growth scenario can be posited as a
possible learning mechanism.

Figures \ref{fig:corrProd} and \ref{fig:corrComp} show how the age of
production and comprehension, respectively, correlates with the degree
(or indegree for the semantic networks). For ease of visual comparison,
the predictor (i.e., the degree) was centered and scaled across
languages. The plots show, overall, a negative correlation between the
month of acquisition and the degree. In production data, the average
correlation across languages was -0.24 (SD=0.10) for the semantic
networks and -0.30 (SD=0.05) for the phonological networks. In
comprehension data, the average correlation was -0.21 (SD=0.08) for the
semantic networks and -0.21 (SD=0.07) for the phonological networks.
These results indicate that nouns with higher degrees are generally
learned earlier, thus replicating previous findings in English (e.g.,
Storkel 2004, 2009; Hills et al. 2009) and extending these findings to
10 different languages, generally, in both production- and
comprehension-based vocabularies.

@@

Overall, we could not reject the null hypothesis of a power-law
distribution: The \(p\)-value was generally above 0.1 in almost all
languages for both production and comprehension. That said, phonological
networks had relatively larger cut-offs than semantic networks. As was
suggested by Arbesman et al. (2010), these \enquote{truncated}
power-laws in phonological networks may reflect the various constraints
that exist on word formation in the phonological domain such as the
number of phonemes in the language, the phonotactics (i.e., the way
sound sequences are arranged in words), and the length of words. Such
constraints may limit the number of words that are phonologically
similar, thus leading to distributions which decay faster than a
non-truncated power law.

@@

We estimated the parameter \(\beta\) using a Bayesian approach. The
inference was performed using the probabilistic programming language
WebPPL (N. Goodman \& Stuhlmuller, 2014). We defined a uniform prior
over \(\beta\), and at each month, we computed the likelihood function
over words that could possibly enter the lexicon at that month, fit to
the words that have been learned at that month (using formula 1). Markov
Chain Monte Carlo sampling resulted in a posterior distribution over
\(\beta\), which we summarized in Figure \ref{fig:growthPred}.

First, the results replicate Hills et al.'s original finding regarding
the semantic network in English and the production-based vocabulary,
which is that this network grows by EXT, not by INT. Second, our results
show that, generally speaking, this finding generalizes to
comprehension-based vocabulary, and holds across languages. This
generalization was obtained in both the semantic\footnote{One could
  imagine that the fact of using English free association norms
  cross-linguistically would decrease the effect of non-English semantic
  networks because of possible cultural differences. However, our
  findings do not support this assumption, rather it supports our
  initial approximation about the universality of the semantic
  similarity measure. That said, this apprimation is not perfect; for
  example there is evidence that a small part of the variance in free
  assication data can be explained by phonological similarity
  (Kachergis, Cox, \& Jones, 2011; Matusevych \& Stevenson, 2018), thus
  leading to possibly minor cross-linguistic differences.} and
phonological domains.

@@

Since we found INT to be uninformative, we dropped it from this
analysis, keeping only EXT. This simplified the model because we no
longer needed to fit growth month-by-month. The latter was a requirement
only for INT where the words' utilities varied from month to month,
depending on how connectivity changed in the growing internal network. A
more direct way to assess and compare the contribution of EXT in
relation to other word-level factors is through conducting linear
regressions, where connectivity in the learning environment, frequency
and length predict the age of acquisition.

For word length, we counted the number of phonemes in our generated IPA
transcription. For word frequency, we used the frequency estimates from
Braginsky et al. (2016) where unigram counts were derived based on
CHILDES corpora in each language (MacWhinney, 2014). For each word,
counts included words that shared the same stem (e.g., \enquote{cats}
counts as \enquote{cat}), or words that were synonymous (e.g.
\enquote{father} counts as \enquote{daddy}). Although these frequency
counts use transcripts from independent sets of children, they are based
on large samples, and this allows us to average out possible differences
between children and the specificities of their input (see J. C. Goodman
et al., 2008 for a similar research strategy).

@@

The findings for the new predictors were as follows. Overall, frequency
is the largest and most consistent predictor of age of acquisition in
both comprehension and production data and across languages, endorsing
results for nouns across a variety of analyses (Braginsky et al., 2016;
J. C. Goodman et al., 2008; B. C. Roy et al., 2015). Word length is more
predictive for production than comprehension (and this difference is
very clear in the global model), replicating previous work (Braginsky et
al., under review). Thus, word length seems to reflect the effects of
production's constraints rather than than comprehension's constraints,
i.e., longer words are harder to articulate but they may not be
significantly more difficult to store and access.

>>>>

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Noun Learning as Network Growth: A Cross-linguistic Analysis},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

% Essential manuscript parts
  \title{Noun Learning as Network Growth: A Cross-linguistic Analysis}

  \shorttitle{Noun Learning as Network Growth}

@@

  \abstract{Children tend to produce words earlier when they are connected to a
variety of other words along the phonological and semantic dimensions.
Though this connectivity effect has been extensively documented, little
is known about the underlying developmental mechanism. One view suggests
that learning is primarily predicted by a network growth model where
highly connected words in the child's early lexicon enable learning of
similar words. Another view suggests that learning is driven by highly
connected words in the external learning environment, instead of highly
connected words in the early internal lexicon. The present study tests
both scenarios systematically in both the phonological and semantic
domains across 10 languages. We show that external connectivity in the
learning environment drives growth in both production- and
comprehension-based vocabularies, even controlling for word frequency
and length. This pattern of findings suggests a word learning mechanism
where children harness their statistical learning abilities to detect
and learn highly connected words in the learning environment.}
  \keywords{Word learning; semantic network; phonological network; network growth;
cross-linguistic analysis. \\

@@

Besides word-level properties, the structure of the lexicon (that is,
how words relate to one another) was also found to predict the Age of
Acquisition (AoA) of words. The lexical structure can be characterized
in terms of a network where each node represents a word in the
vocabulary, and each link between two nodes represents a relationship
between the corresponding pair of words (e.g., Collins \& Loftus, 1975;
Luce \& Pisoni, 1998). Previous studies have investigated early
vocabulary structure by constructing networks using a variety of
word-word relations including shared semantic features (McRae, Cree,
Seidenberg, \& McNorgan, 2005), target-cue relationships in free
association norms (Nelson, McEvoy, \& Schreiber, 1998), co-occurrence in
child directed speech (MacWhinney, 2014), and phonological relatedness
(Vitevitch, 2008). These studies have generally found that children tend
to produce nouns that have higher neighborhood density (i.e., high
connectivity in the network) earlier, both at the phonological and the
semantic level (Carlson, Sonderegger, \& Bane, 2014; Hills, Maouene,
Riordan, \& Smith, 2010; Hills, Maouene, Maouene, Sheya, \& Smith, 2009;
Stella, Beckage, \& Brede, 2017; Storkel, 2009).

While most studies have focused on the static properties of the lexical
network, a few have investigated the underlying developmental process.
In particular, Steyvers and Tenenbaum (2005) suggested that the observed
effects of connectivity are the consequence of how the lexical network
gets constructed in the child's mind. According to this explanation,
known as Preferential Attachment, highly connected words in the child's
lexicon tend to \enquote{attract} more words over time, in a
rich-get-richer scenario (Barabasi \& Albert, 1999). In other words,
what predicts learning is the \emph{internal} connectivity in the
child's early lexicon. In contrast, Hills et al. (2009) suggested that
what biases the learning is not the connectivity in the child's internal
lexicon but, rather, \emph{external} connectivity in the learning
environment. They called this alternative explanation Preferential
Acquisition. For clarity of reading, we will call preferential
attachment the Internally-driven mechanism (INT), and preferential
acquisition the Externally-driven mechanism (EXT).
Figure~\ref{fig:growth} shows an illustration of both growth scenarios
with the same simplified network.

These two proposals represent two divergent ideas about the role of
lexical networks in acquisition. On the INT proposal, learning is driven
by known words with high connectivity to other known words (Figure 1,
left). Thus, the network structure is a causal factor in word learning,
that is, children rely on the organization of their past knowledge to
determine future learning (Altvater-Mackensen \& Mani, 2013; Borovsky,
Ellis, Evans, \& Elman, 2016; Chi \& Koeske, 1983; Storkel, 2009). In
contrast, on the EXT approach, learning is driven by the connectivity of
words that are not known yet (Figure 1, right). Thus, the relevant
network structure is not internally represented by children, and the
observed connectivity effect might be an epiphenomenon of some
properties of the linguistic input. For example, highly connected
concrete nouns in the input could be more easily learned because of
their contextual diversity, allowing for an easier meaning
disambiguation (McMurray, Horst, \& Samuelson, 2012; Smith \& Yu, 2008;
Yurovsky \& Frank, 2015). Another reason could be that these words are
emphasized by the caregivers in their child-directed speech (Clark,
2007; Hoff \& Naigles, 2002; Huttenlocher et al., 1991).

@@

Hills et al. (2009) investigated the growth of lexico-semantic networks
in toddlers and found that growth did not proceed according to INT as
was originally hypothesized by Steyvers and Tenenbaum (2005), but rather
according to
EXT.\footnote{Besides INT and EXT, the authors tested a third mechanism (called the lure of associates) which resembles EXT in that it is driven by the connectivity of external nodes, except that this connectivity is computed with respect to words that are known. However, EXT is the externally-driven scenario that best predicted the data in this previous work.}
This is an important finding because it suggests that learning in the
early stages is mostly driven by properties of the external input,
regardless of how past knowledge is organized. However, this work has
been done in the special case of networks that are based on 1) semantic
associations, 2) production-based vocabularies, and 3) data from
English-learning children, only. The extent to which this result depends
on the domain (e.g., semantic vs.~phonological connectivity), the
vocabulary measure (production vs.~comprehension) and culture/language
is still unclear. In this work, we test the generality of the finding
along these three dimensions.

First, we study the phonological network in addition to the semantic
network. These two networks represent different ways the mental lexicon
is structured (Beckage \& Colunga, 2016). In particular, words that are
neighbors in the semantic network (e.g., \enquote{cat}, \enquote{dog})
are not necessarily neighbors in the phonological network, and vice
versa. Does the phonological network also predict word learning?
Previous work did find an effect of nouns' connectivity in the
phonological network on their age of learning (Carlson et al., 2014;
Stella et al., 2017; Storkel, 2009). In other words, nouns learned
earlier in life tend to sound similar to many other nouns than a word
learned later in life. However, this finding is \emph{a priori}
compatible with both INT and EXT, and previous studies did not
explicitly compare these two mechanisms. Here, we investigate whether
phonological networks, like semantic networks, grow through EXT, or if
they rather grow via INT (Figure~\ref{fig:growth}).

Second, we study vocabularies measured using both comprehension and
production. Previous studies have found differences between these
vocabularies in terms of their content and rate of acquisition (Bates,
Dale, \& Thal, 1995; Benedict, 1979; Fenson et al., 1994). These
differences may reflect the fact that comprehension and production do
not share the same constraints. For instance, whereas comprehension
depends on the ease with which words are stored and accessed, production
depends, additionally, on the ease with which words are articulated,
e.g., shorter words are produced earlier (Braginsky et al., 2016). By
investigating comprehension-based vocabularies, we assess the extent to
which the network growth mechanism captures general learning patterns
beyond the specific constraints of production.

@@

We adopted the following research strategy. We used a cross-linguistic
dataset of children's vocabulary development (Fenson et al., 1994;
Frank, Braginsky, Yurovsky, \& Marchman, 2017), and we studied the
timeline of noun learning using the normative age of acquisition (i.e.,
the age at which at least 50\% of children know a given words). Our
choice of studying the normative learning trajectory rather than the
individual trajectories was motivated by the nature of the dataset
used---which is primarily based on cross-sectional studies. Children may
vary in their individual learning trajectories, but the aggregate data
provide highly robust measures of the average learning patterns (Fenson
et al., 1994). The use of such measures has lead to important insights
on the mechanisms of word learning (Braginsky et al., 2016; J. C.
Goodman et al., 2008; Hills et al., 2010, 2009; Stella et al., 2017;
Storkel, 2009).

The paper is organized as follows. First, we describe the datasets we
used and explain how we constructed the networks. Second, we analyze
static properties of words' connectivity in these networks (correlation
with age of acquisition and shape of the distribution) and we explain
how these properties inform hypotheses about network growth. Next, we
explicitly fit the two hypothesized growth mechanisms to the data. We
investigate the extent to which the results obtained in Hills et al.
(2009) generalize to phonological networks and comprehension-based
vocabularies, and whether this generalization holds
cross-linguistically.

@@

We used data from Wordbank (Frank et al., 2017), an open repository
aggregating cross-linguistic language developmental data of the
MacArthur-Bates Communicative Development Inventory (CDI), a parent
report vocabulary checklist. Parent report is a reliable and valid
measure of children's vocabulary that allows for the cost-effective
collection of datasets large enough to test network-based models of
acquisition (Fenson et al., 1994). When filling out a CDI form,
caregivers are either invited to indicate whether their child
\enquote{understands} (comprehension) or \enquote{understands and says}
(production) each of about 400-700 words. For younger children (e.g., 8
to 18 months in the English data), both comprehension and production are
queried, whereas for older children (16 to 36 months) only production is
queried. Due to these limitations, we use data from younger children to
test comprehension and data from older children to test production. In
addition, following previous studies (Hills et al., 2009; Storkel,
2009), we restricted our analysis to the category of nouns due to the
fact that nouns predominate the early expressive and receptive lexicons
(Bates et al., 1995). Their larger sample size (compared, for example,
to verbs or adjectives) is more suited to the network-based analysis of
development. Table \ref{tab:stats} gives an overview of the data we
used.

@@

We constructed semantic networks for English data following the
procedure outlined in Hills et al. (2009), as follows. We used as an
index of semantic relatedness the Florida Free Association Norms (Nelson
et al., 1998). This dataset was collected by giving adult participants a
word (the cue), and asking them to write the first word that comes to
mind (the target). For example, when given the word \enquote{ball}, they
might answer with the word \enquote{game}. A pair of nodes were
connected by a directed link from the cue to the target if there was a
cue-target relationship between these nodes in the association norms.
The connectivity of a given node was characterized by its
\emph{indegree}: the number of links for which the word was the
target.\footnote{This choice was based on prior work by Hills et al. (2009) stating that analyses with both outdegrees (sum of the links where the word is the cue in a cue-target pair) and total degree  (outdegree plus indegree) led to results weaker than those calculated with indegree.}
To model growth from month to month, we constructed a different network
at each month, based on the nouns that have been acquired by that month.

Since the free association norms are available only in English, we used
the hand-checked translation equivalents available in Wordbank, which
allowed us to use the English association norms across 10 languages.
Using the same association data across languages does not necessarily
lead to similar networks. Indeed, though this approximation assumes that
the semantic similarity measure is universal---which is a reasonable
approximation (e.g., Youn et al., 2016)---the set of words acquired by
children as well as the timeline of this acquisition can still vary from
language to language leading to possibility different learning
strategies.

@@

To construct phonological networks we first mapped the orthographic
transcription of words to their International Phonetic Alphabet (IPA)
transcriptions in each language, using the open source text-to-speech
software \textbf{\href{http://http://espeak.sourceforge.net/}{Espeak}}.
This software provides the correct IPA transcription if the word is
found in a spelling-to-phonemes dictionary, otherwise it uses
language-specific pronunciation rules to generate an approximate
phonetic transcription. We used the Levenshtein distance (also known as
edit distance) as a measure of phonological relatedness between two
nodes. The measure counts the minimum number of operations (insertions,
deletions, substitutions) required to change one string into another.

In previous studies, two nodes were linked if they had an edit distance
of 1 (Carlson et al., 2014; Stella et al., 2017; Storkel, 2009).
However, these studies reported a contribution of phonological
connectivity to noun learning when networks were built using a rich
adult vocabulary. Since the focus of the current study is on the
mechanism of growth, the networks should be based on children's early
vocabulary. The latter, however, contains very few noun pairs with an
edit distance of 1. Thus, we increased the threshold from 1 to 2, that
is, two nodes were related if their edit distance was equal to 1 or
2.\footnote{We also considered the case of an edit distance of 1 as well as the continuous measure, i.e., the inverse edit distance without threshold. In both cases, the results were weaker than those obtained with a threshold of 2. We did not consider the case of a threshold larger than 2 since many pairs become phonologically unrelated.}
The connectivity of a given node was characterized with its
\emph{degree}: the number of links it shares with other words.

@@

We start by analyzing word connectivity in the global (static) network.
We constructed this network using nouns learned by the oldest age for
which we have CDI data (e.g., in English this corresponds, in
comprehension, to the network by 18 months, and in production, to the
network by 30 months). This global network is the end-state towards
which both INT and EXT converge by the last month of learning. Moreover,
following Hills et al. (2009), we used this end-state network as a proxy
for the external connectivity in the learning environment. Below we
analyze properties of this global networks that may a priori hint at an
INT- or EXT-like growth. In order to compare various predictors on the
same data, we restrict the analysis to the subset of nouns for which we
had semantic associates in the free association norms.

@@

Connectivity in the global network is directly related to EXT as it
represents the explicit criterion this growth scenario uses to determine
what words should be learned first (Figure \ref{fig:growth}). Therefore,
a direct consequence of an EXT-like growth scenario is a correlation
between connectivity in the global network and the age of acquisition.
This correlation is also necessary to INT, although the causality is
reversed: Higher connectivity in the global network is caused by earlier
learning, not the other way around. Some words end up being highly
connected in the global network precisely because they happen to be
acquired earlier and, therefore, have a higher chance of accumulating
more links over time. Thus, the correlation between connectivity in the
end-state network and AoA can result from both EXT and INT. If there is
no such correlation, neither growth scenario can be posited as a
possible learning mechanism.

Figures \ref{fig:corrProd} and \ref{fig:corrComp} show how the age of
production and comprehension, respectively, correlates with the degree
(or indegree for the semantic networks). For ease of visual comparison,
the predictor (i.e., the degree) was centered and scaled. The plots
show, overall, a negative correlation between the month of acquisition
and the degree. In production data, the average correlation across
languages was -0.24
(\$SD=\(0.10) for the semantic networks and -0.30 (\)SD=\(0.05) for the phonological networks. In comprehension data, the average correlation was -0.21 (\)SD=\(0.08) for the semantic networks and -0.21 (\)SD=\$0.07)
for the phonological networks. These results indicate that nouns with
higher degrees are generally learned earlier, thus replicating previous
findings in English (Hills et al., 2009; Storkel, 2009) and extending
these findings to 10 different languages, generally, in both production-
and comprehension-based vocabularies.

@@

Overall, we could not reject the null hypothesis of a power-law
distribution: The \(p\)-value was generally above 0.1 in almost all
languages for both production and comprehension. That said, phonological
networks had relatively larger cut-offs than semantic networks. These
\enquote{truncated} power-laws in phonological networks may be due to
the constraints that exist on word formation in the phonological domain
such as the size of the phonemic inventory, phonotactic rules, and word
length. Such constraints may limit the number of words that are
phonologically similar, thus leading to distributions which decay faster
than a non-truncated power law (Arbesman et al., 2010).

@@

We estimated the parameter \(\beta\) using a Bayesian approach. The
inference was performed using the probabilistic programming language
WebPPL (N. Goodman \& Stuhlmuller, 2014). We defined a uniform prior
over \(\beta\), and at each month, we computed the likelihood function
over words that could possibly enter the lexicon at that month, fit to
the words that have been learned at that month (using formula 1). Markov
Chain Monte Carlo sampling resulted in a posterior distribution over
\(\beta\), which we summarized in Figure \ref{fig:growthPred}. The
results replicate Hills et al.'s original finding regarding the semantic
network in English and the production-based vocabulary, which is that
this network grows by EXT, not by INT. Crucially, our results show that,
generally speaking, this finding generalizes to comprehension-based
vocabulary, and holds across languages. This generalization was obtained
in both the semantic\footnote{One could imagine that the fact of using
  English free association norms cross-linguistically would decrease the
  effect of non-English semantic networks because of possible cultural
  differences. However, our findings do not support this assumption,
  rather, it supports our initial approximation about the universality
  of the semantic similarity measure. That said, this approximation is
  not perfect. For example there is evidence that a small part of the
  variance in free association data can be explained by phonological
  similarity (Kachergis, Cox, \& Jones, 2011; Matusevych \& Stevenson,
  2018), thus leading to possibly minor cross-linguistic differences.}
and phonological domains.

@@

Since we found INT to be uninformative, we dropped it from this
analysis, keeping only EXT. This simplified the model because we no
longer needed to fit growth month-by-month. The latter was a requirement
only for INT where the words' utilities varied from month to month,
depending on how connectivity changed in the growing internal network. A
more direct way to assess and compare the contribution of EXT in
relation to other word-level factors is through conducting regressions,
where connectivity in the learning environment, frequency and length
predict the age of acquisition.

For word length, we counted the number of phonemes in our generated IPA
transcription. For word frequency, we used the frequency estimates from
Braginsky et al. (2016) where unigram counts were derived based on
CHILDES corpora in each language (MacWhinney, 2014). Although these
frequency counts use transcripts from independent sets of children, they
are based on large samples, and this allows us to average out possible
differences between children and the specificities of their input (see
J. C. Goodman et al., 2008 for a similar research strategy).

@@

The findings for the new predictors were as follows. Overall, frequency
is the largest and most consistent predictor of age of acquisition in
both comprehension and production data and across languages, endorsing
results for nouns across a variety of analyses (Braginsky et al., 2016;
J. C. Goodman et al., 2008; B. C. Roy et al., 2015). Word length is more
predictive for production than comprehension (and this difference is
very clear in the global model), replicating previous work (Braginsky et
al., under review). Thus, word length seems to reflect the effects of
production's constraints rather than comprehension's constraints, i.e.,
longer words are harder to articulate but they may not be significantly
more difficult to store and access.

As for the factors of interest, i.e., semantic and phonological
connectivity, we found cross-linguistic differences. Connectivity
contributes to learning in some languages but not in other. In
particular, semantic connectivity does not explain variance in English
data beyond that explained by phonological connectivity, frequency and
length. This finding contrasts with the original finding in Hills et al.
(2009). However, this difference might be due to our using a slightly
different model (which included word length as a covariate) and a larger
dataset. That said, and despite these apparent cross-linguistic
differences, both phonological and semantic connectivity are significant
predictors in the combined model (Figure \ref{fig:staticAll}).
